Gradual Soundness: Lessons from Static Python

### 2021-10-27 ecoop-2022/paper.tex
ecoop: type Dyn notes

\begin{itemize}
  \item
    Evidence that gradual soundness delivers fast performance incrementally,
    as programmers migrate to sound types.
    \SP{} can be applied to a codebase with no refactoring and no performance overhead.
    After refactoring to sound types, \SP{} performs well on a microbenchmark
    suite and in production.
  \item
    A formal description and analysis of the \SP{} core language.
    The analysis includes a proof of type soundness,
    a proof that optimization soundly improves code,
    and [FILL perhaps] a metric to quantify the refactoring cost and optimization benefit
    of a type migration.
  \item
    A list of the expressiveness limitations in \SP{} relative to
    idealized gradual type systems,
    and an informal description of the programming discipline that Instagram
    follows in their gradually-sound codebase.
\end{itemize}

@@

@@

  \item \textbf{No Overloading}.
    Methods cannot be overloaded; each combination of a class
    and a name can refer to at most one method.
    This ensures that method resolution depends only on the
    type of the receiver object.
    % resolution does NOT depend on result type (T doesn't matter in `x:T = o.m()`)
    % Q. does resolution depend on args?

>>>>

\begin{itemize}
  \item
    Evidence that gradual soundness delivers fast performance incrementally,
    as programmers migrate to sound types.
    \SP{} can be applied to a codebase with no refactoring and no performance overhead.
    After refactoring to sound types, \SP{} performs well on a microbenchmark
    suite and in production.
  \item
    A formal description and analysis of the \SP{} core language.
    The analysis includes a proof of type soundness,
    a proof that optimization soundly improves code,
    and [FILL perhaps] a metric to quantify the refactoring cost and optimization benefit
    of a type migration.
  \item
    In-depth discussions of the expressiveness limitations in \SP{} relative to
    idealized gradual type systems, and the programming discipline that Instagram
    follows in their gradually-sound codebase.
\end{itemize}

@@

No way to write the dynamic type.
Instead, it fills in for absent annotations.
Either totally missing or a missing generic parameter.

%% TODO are there any types C[T0, T1] where you might want C[dyn, T1]? how to express?

Cannot appear everywhere ... ChkDict can't hold a Dynamic,
and primitives are incompatible with Dynamic.

Dynamic behavior is always (?) the same as normal Python.

@@

  \item \textbf{No Overloading}.
    Methods cannot be overloaded. Each combination of a class
    and a name can refer to at most one method.
    This ensures that method resolution depends only on the
    type of the receiver object.
    % resolution does NOT depend on result type (T doesn't matter in `x:T = o.m()`)
    % Q. does resolution depend on args?
